\documentclass{article}
\usepackage{3Rdefs}

\title{Understanding Reality Through Logic: Introducing 3R}
\date{}

\begin{document}

\maketitle

\section*{Language, Logic, and Uncertainty}

\subsection*{The Nature of Language}

Human language operates in two domains: the real and the imaginary. The real domain is rooted in our brain's visual and sensory systems, which model objects amazingly well. Naming these objects enables person to person communication about real objects they observe. However, language also includes imaginary words, terms without sensory foundations, often leading to ambiguity, inconsistency, and foolish arguments.

Imaginary words are great for fiction, but in the form of logic, they are also essential for logical modelling. Logic is an exception to the imaginary word rule. You are not free to just invent logical words. Logical words or ideally symbols must be consistently formed from the base definition of 0 as NOT 1. Logic is a work of fiction, independent of reality but it is an amazingly powerful one due to its guaranteed true-false consistency. Logic defines true and false—these are imaginary, not real words, however, we must have very strict rules on how these logical words are formed—they must not be allowed to be political (axiomatic or belief-based).

A critical issue arises from illogically defined word combinations that blur the distinction between consistently defined logical terms and those describing sensory models. For instance, the term "real number" fails to understand both reality and logic. This term leads to other highly illogical combinations such as "countable infinity" and "infinite set." The term "real number" is arguably one of the most illogical word combinations ever, leading to scientific and logical confusion. However, non-rational numbers are highly logical. There is no requirement to restrict logic to only those numbers defined by the division operator. Mathematicians and scientists must develop a clearer understanding of the distinction between consistently defined logical terms and those describing sensory models.

\subsection*{The Role of Pure Logic and Modelling Uncertainty}

Pure logic, based on binary values (0 defined as NOT 1), eliminates ambiguity by disallowing contradictions. Logic is foundational to mathematics and computing, providing a precise language for scientific models. However, reality, while scientifically modelled by 0 and 1, requires the definition of \qbit{} to incorporate uncertainty in our models. The concept of \qbit{} represents the superposition of discrete and continuous states, embodying both chance and causality. By incorporating \qbit{}, logical models can better reflect the complexities of reality.

\subsection*{The Challenge of Infinity}

Infinity, or \(\infty\), is a logically defined term, not a real object. It is no more real than 0 or 1. Traditional binary logic struggles to represent infinite sequences or continuous processes fully, leading to confusion among mathematicians and scientists. By defining \qbit{}, we create a framework that accommodates the endless progression of sequences and the continuous nature of reality without falling into logical contradictions. The diagonal argument shows we can construct an infinite list of ever more precise values to count the \([0,1]\) continuum, acknowledging that our models are always expanding but never complete.

\subsection*{Three Relativistic Dimensions (\iR{})}

Three Relativistic Dimensions (\iR{}) is the recognition that logical models require three relativistic dimensions to avoid binary contradictions caused by independent binary dimensions. Unlike traditional models that rely on independent dimensions (length, width, height, time), \iR{} uses three interconnected relativistic dimensions. Evolution has optimised sensory perception using three dimensions because it is the best possible model. This approach does not imply that the universe is constructed from logic; indeed, logic can prove that it is not. Rather, \iR{} forms the best possible, most logical description of the universe. While many scientific models exceed human capabilities in specific areas, the concept of three relativistic dimensions is logically fundamental to the most complete and accurate models, such as the three-dimensional neural models animal brains construct from sensory inputs. The relativistic sphere, closely related to a three-dimensional sphere, is based on relativistic, not independent, dimensions, offering a more comprehensive understanding of reality.

\subsection*{Incompleteness and Certainty}

Logical models must always acknowledge their inherent incompleteness and declare their limits of certainty. While models can achieve exceptionally high certainty—such as the three-dimensional neural models animal brains construct from sensory inputs—they are never perfect. The recognition of \iR{}, however, offers the best possible representation of reality.

\subsection*{Conclusion}

Pure logic, based on binary values (0 defined as NOT 1) and incorporating the concept of \qbit{}, provides the only non-contradictory foundation for scientific models. By recognising the limitations and declaring the certainty of our models, we can achieve a highly accurate understanding of reality. Recognising the necessity of three relativistic dimensions (\iR{}) offers the most comprehensive and precise framework, integrating the certainty of logic with the complexities of the world. By embracing \qbit{}, we can address the challenges of infinity and create models that reflect both continuous and discrete aspects of nature and the universe. The concept of three relativistic dimensions is logically fundamental to the most complete and accurate models, guiding us towards a more profound understanding of reality.

\end{document}


